---
title: "Assignment 9/11 forecast"
output: html_notebook
---
```{r}
Sept11Travel.data<-read.csv("D:\\BAD\\Sept11Travel.csv")
Sept11Travel.data

```


Sept11Travel.ts<-ts(Sept11Travel.data, start = c(1990, 1), end = c(2004, 4), freq = 12)
preSept11Travel.ts <- window(Sept11Travel.ts, start = c(1990, 1), end = c(2001, 9))
postSept11Travel.ts <- window(Sept11Travel.ts, start = c(2001,9), end = c(2004,4))

library("forecast")
```

```{r}
Air.ts<-ts(Sept11Travel.data$Air.RPM..000s. , start = c(1990, 1), end = c(2004, 4), freq = 12)
preSept11Air.ts <- window(Air.ts, start = c(1990, 1), end = c(2001, 8))
postSept11Air.ts <- window(Air.ts, start = c(2001,9), end = c(2004,4))
par("mar")
par(mar=c(1,1,1,1))

plot(preSept11Air.ts, xlab = "Time", ylab = "Air Travel (in 000s)", ylim=c(29672400,69003600))
#Increase in Air travel pre Sept 11

```

```{r}

#Adjusting seasonality multiplicatively
require(graphics)
decompose_preSept11Air.ts = decompose(preSept11Air.ts, "multiplicative")
adjust_preSept11Air.ts = preSept11Air.ts/decompose_preSept11Air.ts$seasonal
plot(adjust_preSept11Air.ts, col = "Blue",lwd = 2)


#checking the best model
adjust_preSept11.Air.lm<-tslm(adjust_preSept11Air.ts ~ trend)
summary(adjust_preSept11.Air.lm)

adjust_preSept11.Air.lm.season<-tslm(adjust_preSept11Air.ts ~ season)
summary(adjust_preSept11.Air.lm.season)

adjust_preSept11.Air.lm.seasonalityTrend<-tslm(adjust_preSept11Air.ts ~ trend + season)
summary(adjust_preSept11.Air.lm.seasonalityTrend)

#residuals
adjust_preSept11.Air.lm.seasonalityTrend$residuals
```

```{r}
#ACF for residuals
Acf(adjust_preSept11.Air.lm.seasonalityTrend$residuals, lag.max = 12, main="")
```

```{r}
#Model on Air with linear trend and addtive seasonality
preSept11Air.lm<- tslm(preSept11Air.ts ~ trend )
preSept11Air.lm
#Adding season
preSept11Air.lm.season <- tslm(preSept11Air.ts ~ trend + season)
summary(preSept11Air.lm.season)
par("mar")
par(mar=c(1,1,1,1))
plot(preSept11Air.lm.season.ts)

lines(preSept11Air.lm$fitted, lwd = 1)
#Adjusting seasonality additive
require(graphics)
decomposed_preSept11Air.ts = decompose(preSept11Air.ts, "additive")
adjusted_preSept11Air.ts = preSept11Air.ts-decomposed_preSept11Air.ts$seasonal
plot(adjusted_preSept11Air.ts, col = "Blue",lwd = 2)
adjusted_preSept11.Air.lm.Trend<-tslm(adjusted_preSept11Air.ts ~ trend)
summary(adjusted_preSept11.Air.lm.Trend)

Airnaive.pred <- forecast::naive(adjusted_preSept11Air.ts, h = 36)
Airsnaive.pred <- forecast::snaive(adjusted_preSept11Air.ts, h = 36)
Airnaive.pred
Airsnaive.pred
```

```{r}
#forecast
plot(Air.ts,  ylab = "Air Passengers Revenue Miles ($BillionS)", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2005), main = "")

axis(1, at = seq(1990, 2005, 1), labels = format(seq(1990, 2005, 1)))
lines(Airnaive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(Airsnaive.pred$mean, lwd = 2, col = "red", lty = 1)
lines(c(2004.70 - 3, 2004.70 - 3), c(0, 700999930))
lines(c(2004.25, 2004.25), c(0, 700999930))
text(1994.25, 69103630, "Training")
text(2002.75, 69999990, "Validation")
text(2004.90, 69103630, "Future")
arrows(2001, 67000000, 1990, 67000000, code = 3, length = 0.1, lwd = 1,angle = 30)
```



```{r}
#Rail ts
Rail.ts<-ts(Sept11Travel.data$Rail.PM , start = c(1990, 1), end = c(2004, 4), freq = 12)
preSept11Rail.ts <- window(Rail.ts, start = c(1990, 1), end = c(2001, 8))
postSept11Rail.ts <- window(Rail.ts, start = c(2001,9), end = c(2004,4))

plot(preSept11Rail.ts, xlab = "Time", ylab = "Rail Travel (in 000s)")
#additive seasonality
decomposed_preSept11Rail.ts = decompose(preSept11Rail.ts, "additive")
adjusted_preSept11Rail.ts = preSept11Rail.ts-decomposed_preSept11Rail.ts$seasonal
plot(adjusted_preSept11Rail.ts, col = "Blue",lwd = 2)
adjusted_preSept11.Rail.lm.Trend<-tslm(adjusted_preSept11Rail.ts ~ trend)
summary(adjusted_preSept11.Rail.lm.Trend)

adjusted_preSept11.Rail.lm.Trendquad<- tslm(adjusted_preSept11Rail.ts ~ trend + I(trend^2))
summary(adjusted_preSept11.Rail.lm.Trendquad)

Railnaive.pred <- forecast::naive(adjusted_preSept11Rail.ts, h = 36)
Railsnaive.pred <- forecast::snaive(adjusted_preSept11Rail.ts, h = 36)
Railnaive.pred
Railsnaive.pred

#forecast
plot(Rail.ts,  ylab = "Rail Passengers", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2005), main = "")

axis(1, at = seq(1990, 2005, 1), labels = format(seq(1990, 2005, 1)))
lines(Railnaive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(Railsnaive.pred$mean, lwd = 2, col = "red", lty = 1)
lines(c(2004.70 - 3, 2004.70 - 3), c(0, 700999930))
lines(c(2004.25, 2004.25), c(0, 700999930))
text(1994.25, 69100000, "Training")
text(2002.75, 69990000, "Validation")
text(2004.90, 69100000, "Future")
arrows(2001, 67000000, 1990, 67000000, code = 3, length = 0.1, lwd = 1,angle = 30)
```

```{r}
Car.ts<-ts(Sept11Travel.data$VMT..billions. , start = c(1990, 1), end = c(2004, 4), freq = 12)
preSept11Car.ts <- window(Car.ts, start = c(1990, 1), end = c(2001, 8))
postSept11Car.ts <- window(Car.ts, start = c(2001,9), end = c(2004,4))

plot(preSept11Car.ts, xlab = "Time", ylab = "Car Travel (in 000s)")
#adjusting seasonality additively
decomposed_preSept11Car.ts = decompose(preSept11Car.ts, "additive")
adjusted_preSept11Car.ts = preSept11Car.ts-decomposed_preSept11Car.ts$seasonal
plot(adjusted_preSept11Car.ts, col = "Blue",lwd = 2)
adjusted_preSept11.Car.lm.Trend<-tslm(adjusted_preSept11Car.ts ~ trend)
summary(adjusted_preSept11.Car.lm.Trend)

Carnaive.pred <- forecast::naive(adjusted_preSept11Car.ts, h = 36)
Carsnaive.pred <- forecast::snaive(adjusted_preSept11Car.ts, h = 36)
Carnaive.pred
Carsnaive.pred
#forecast
plot(Car.ts,  ylab = "Vehicle Travel", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2005), main = "")

axis(1, at = seq(1990, 2005, 1), labels = format(seq(1990, 2005, 1)))
lines(Carnaive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(Carsnaive.pred$mean, lwd = 2, col = "red", lty = 1)
lines(c(2004.70 - 3, 2004.70 - 3), c(0, 700999930))
lines(c(2004.25, 2004.25), c(0, 700999930))
text(1994.25, 69103630, "Training")
text(2002.75, 69999990, "Validation")
text(2004.90, 69103630, "Future")
arrows(2001, 67000000, 1990, 67000000, code = 3, length = 0.1, lwd = 1,angle = 30)
```

                  



